Group Fou\\*R* Term Project - Summary of Raw Data, Explanations and Preliminary Analysis

### 2018-12-16 Draft.Rmd
.

We worked on trade statistics data of Turkey against the G7 countries and 3 more countries (Russia, India, China) both for export and import.Our aim is to analyze Turkey’s trade characteristics by product groups. Within this aim, we take trade values and quantities by product groups in MS Excel file format from trademap.org website for each trade partner country for the period of 2013 - 2017. 
In this study, We applied data manipulation to make it proper for analysis and explanatory data analysis with visulations and conclude with overall statistical overview of Turkey's trade.

@@

Following the extraction from source (Trademap.org), our raw data consist of 10 countries’ trade data against Turkey. This data splits to 2 with trade direction as “Export” and “Import”. The other dimension is stat for trade, one is “Quantity” of Export/Import, the other is “Value” in billion USD of Export/Import. Therefore we have four main data frame for each country (total of 40 excel files).

@@

<tbody>
<tr class="odd">
<td align="left">ExportValues</td>
<td align="left">Export from Turkey to target countries and their values in billion USD</td>
</tr>
<tr class="odd">
<td align="left">ExportQuantity</td>
<td align="left">Export from Turkey to target countries and their quantities in units(tons,gallons etc)</td>
</tr>
<tr class="odd">
<td align="left">ImportValues</td>
<td align="left">Import to Turkey from target counties and their values in billion USD</td>
</tr>
<tr class="odd">
<td align="left">ImportQuantity</td>
<td align="left">Import to Turkey from target counties and their quantities in units(tons,gallons etc)</td>
</tr>

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(shiny)
library(tidyverse)
library(dplyr)
library(readxl)
library(leaflet)
```

@@

>>>>

We worked on trade statistics data of Turkey against the G7 countries and 3 more countries (Russia, India, China) both for export and import.Our aim is to analyze Turkey?s trade characteristics by product groups. Within this aim, we take trade values and quantities by product groups in MS Excel file format from trademap.org website for each trade partner country for the period of 2013 - 2017. 
In this study, We applied data manipulation to make it proper for analysis and explanatory data analysis with visulations and conclude with overall statistical overview of Turkey's trade.

@@

Following the extraction from source (Trademap.org), our raw data consist of 10 countries? trade data against Turkey. This data splits to 2 with trade direction as ?Export? and ?Import?. The other dimension is stat for trade, one is ?Quantity? of Export/Import, the other is ?Value? in thousand USD of Export/Import. Therefore we have four main data frame for each country (total of 40 excel files).

@@

<tbody>
<tr class="odd">
<td align="left">ExportValues</td>
<td align="left">Export from Turkey to target countries and their values in thousand USD</td>
</tr>
<tr class="odd">
<td align="left">ExportQuantity</td>
<td align="left">Export from Turkey to target countries and their quantities in units(tons,gallons etc)</td>
</tr>
<tr class="odd">
<td align="left">ImportValues</td>
<td align="left">Import to Turkey from target counties and their values in thousand USD</td>
</tr>
<tr class="odd">
<td align="left">ImportQuantity</td>
<td align="left">Import to Turkey from target counties and their quantities in units(tons,gallons etc)</td>
</tr>

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(shiny)
library(tidyverse)
library(dplyr)
library(readxl)
library(leaflet)
library(reshape2)
```

@@

###Analysis

####SALIH

```{r}
# Export Values in $ 
ttable=export_jTableMELTED%>%group_by(Country,year)%>%
  summarise(co_total=sum(Values,na.rm = TRUE))%>%
  arrange(desc(co_total,year))

```{r}
#Import Values in $
vtable=import_jTableMELTED%>%group_by(Country,year)%>%summarise(co_total=sum(Values,na.rm = TRUE))%>%arrange(desc(co_total,year))

a<-ggplot(data=vtable, aes(x=Country, y=co_total,fill=year)) +
  geom_bar(stat="identity", position=position_dodge(), colour="black")+
  ggtitle("Total Import Values in $ to Selected Countries")

#Export percentage of Countries

b<-ggplot(data=ttable, aes(x=Country, y=co_total,fill=year)) +
  geom_bar(stat="identity", position=position_dodge(), colour="black")+ggtitle("Total Export Values in $ to Selected Countries")

grid.arrange(a,b,  nrow = 2)
```

#Import percentage Values in $

pim_table<-vtable%>%filter(year=='2017')%>%
  mutate(perc_count=co_total/sum((vtable%>%filter(year=='2017'))$co_total)*100)%>%
  select(Country,year,co_total,perc_count)

ggplot(pim_table, aes(x="", y=perc_count, fill=Country)) + geom_bar(stat="identity", width=1)+
  coord_polar("y") + geom_text(aes(label = round(perc_count,1), "%"), position = position_stack(vjust = 0.5))+
  labs(x = NULL, y = NULL, fill = NULL, title = "Percentage of Countries in Import in 2017")+theme_classic()+
  theme(axis.text = element_blank(),plot.title = element_text(hjust = 0.5, color = "#666666"))

####OMER

In order to deeply understand the data; we build up shiny app with Country, Product and Year variables.

Please follow the link for interactive graph: [Link](https://groupfour.shinyapps.io/tradeOrg)

####MUHARREM
####Analysis of Trade Deficit by Countries and Years

To discovery that country's effects to trade deficit, we calculated net_value which formulated as 'Export_value - Import_value'. And visualized that value on the basis of year.

First, group data;

```{r }
export_Product_Group <- export_jTableMELTED%>%
group_by(Country,PcodeTwo,X__2,year)%>%
summarize(Quantity = sum(Quantity), Values = sum(Values))%>%
rename(Product_Group_Code =PcodeTwo, Product_Group_Name = X__2,Export_Quantity = Quantity,Export_Value =Values)

import_Product_Group <- import_jTableMELTED%>%
group_by(Country,PcodeTwo,X__2,year)%>%
summarize(Quantity = sum(Quantity), Values = sum(Values))%>%
rename(Product_Group_Code =PcodeTwo, Product_Group_Name = X__2,Import_Quantity = Quantity,Import_Value =Values)

Export_Import_JTable <- export_Product_Group %>%
  full_join(import_Product_Group,by=c("Country","Product_Group_Code","Product_Group_Name","year")) %>%
mutate(Import_Quantity = replace(Import_Quantity, is.na(Import_Quantity),0),
        Export_Quantity = replace(Export_Quantity, is.na(Export_Quantity),0))%>%
select(Country,Product_Group_Code,Product_Group_Name,year,Export_Quantity,Import_Quantity,Export_Value,Import_Value)
```

Continue with visulation;

```{r}
####Net Value By Country###
Net_Value_By_Country <- Export_Import_JTable %>%
group_by(Country,year)%>%
summarize (Export_Value = sum(Export_Value), Import_Value = sum(Import_Value), 
           Net_Value = sum(Export_Value - Import_Value),Net_value_K = round((sum(Export_Value - Import_Value)/1000),digits=0))

# Total Net Value fiiled with Country on the basis of Year #
ggplot(Net_Value_By_Country, aes(x=year, y=Net_value_K)) + 
  geom_bar(stat='identity', aes(fill=Country), width=.6,position = position_stack(reverse = TRUE))  +
  labs(subtitle="## Total Net Value fiiled with Country on the basis of Year ##", title= "Total Net Values") + 
  coord_flip()+
  theme(legend.position = "top")
```

As seen in the graph, only UK gives trade surplus for all years analysed and China, Russia, Germany and USA have the most significant effect on trade deficit.

##### An Alternative perspective to explore effects of countries

The effects of countries are changing year by year. So that as an alternative perspective we could average net value for all countries in five years and normalized that value to show which country's effects are more than others.

```{r }
#Normalized of Average Values for five years
Normalized_Values_By_Country <- Export_Import_JTable %>%
  #filter (year == 2017)%>%
  group_by(Country)%>%
  summarize (Net_Value = mean(Export_Value - Import_Value))%>%
  mutate(normalized_Net_value =  round((Net_Value  - mean(Net_Value))/sd(Net_Value ),2)) %>%
  mutate(value_Net_type =  ifelse(normalized_Net_value < 0, "below", "above"))

#Diverging Bars for normalized_Export_value
ggplot(Normalized_Values_By_Country, aes(x=Country, y=normalized_Net_value , label=normalized_Net_value)) + 
  geom_bar(stat='identity', aes(fill=value_Net_type), width=.5)  +
  scale_fill_manual(name="Normalised Value", 
                    labels = c("Above Average", "Below Average"), 
                    values = c("above"="seagreen1", "below"="Red")) +  
  labs(subtitle="Normalised Net Value for 5 years by Country'", 
       title= "Diverging Bars") + 
  coord_flip()
```

Considering the 5 years period, it is clear that China, Germany and Russia has contribution to trade deficit above average.

In order to understand the correlation between trade deficit/surplus and BIST100 index and USD/TL rates, we added this economic indicators to our analysis:

Read the USD and BIST100 index values and rename the columns

```{r}
fileUSD= "C:/Users/emrek/Google Drive/BDA/503-EssentialsOfDataAnalytics/GitHub/gpj18-group_four/EVDS.xlsx"
usdR <- read_excel(fileUSD, col_names=FALSE, skip=1)

colnames(usdR) <- c("DateR","ExchangeRate")
usd <- na.omit(usdR)
remove(usdR)

usdN<-usd %>% mutate(Date=dmy(DateR))
usdM<- usdN %>% group_by(year=year(Date)) %>% summarise(Rate=mean(ExchangeRate)*10000000)

fileBIST= "C:/Users/emrek/Google Drive/BDA/503-EssentialsOfDataAnalytics/GitHub/gpj18-group_four/BIST100.xlsx"
bist100R <- read_excel(fileBIST, col_names=FALSE, skip=1)

colnames(bist100R) <- c("Date",	"Year","Value",	"Açılış",	"Yüksek",	"Düşük",	"Hac.",	"Fark %")
bist100 <- bist100R  %>% select(Date,Year,Value) %>% 
  group_by(Year) %>% 
  summarise(Value = mean(Value)*1000)
```

Draw the graph to represent USD/TL rate, BIST100 index and Trade Deficit in time series.

```{r}
sumImport<- import_jTable %>% group_by(Country) %>% summarise(Value_2013=sum(Value_2013),Value_2014=sum(Value_2014),Value_2015=sum(Value_2015),Value_2016=sum(Value_2016),Value_2017=sum(Value_2017))
sumExport<- export_jTable %>% group_by(Country) %>% summarise(Value_2013=sum(Value_2013),Value_2014=sum(Value_2014),Value_2015=sum(Value_2015),Value_2016=sum(Value_2016),Value_2017=sum(Value_2017))
joinedTableDeficit<-sumExport %>%  inner_join(sumImport,by="Country") %>%
  mutate(Value_2013=Value_2013.y-Value_2013.x,Value_2014=Value_2014.y-Value_2014.x, Value_2015=Value_2015.y-Value_2015.x, Value_2016=Value_2016.y-Value_2016.x,
         Value_2017=Value_2017.y-Value_2017.x) %>%
  select(Country, Value_2013, Value_2014, Value_2015, Value_2016, Value_2017)

joinedTableDeficitMelt <- melt(joinedTableDeficit, id = c("Country"))
deficitPerCountry<-joinedTableDeficitMelt %>% mutate(year = case_when( variable == "Value_2013" ~ 2013,
                                                                       variable == "Value_2014" ~ 2014,
                                                                       variable == "Value_2015" ~ 2015,
                                                                       variable == "Value_2016" ~ 2016,
                                                                       variable == "Value_2017" ~ 2017)) %>%
  select(Country,year,value)

deficitPerYear <- deficitPerCountry %>%
  group_by(year) %>% summarise(value=sum(value))

p<-ggplot() + 
  geom_line(data=usdM,aes(x=year, y=Rate,group = 1, colour="USD_TL_Rate"))+
  geom_line(data=deficitPerYear,aes(x=year, y=value,group = 1, colour="Trade_Deficit"))+
  geom_line(data=bist100,aes(x=Year, y=Value,group = 1,colour="BIST100_Index")) +
  scale_colour_manual(name="Indicators",values=c(USD_TL_Rate="red", Trade_Deficit="blue", BIST100_Index="purple"))

print(p)

Trade deficit has negatively correlated with USD/TL rate between 2013 and 2016 as expected and for the year 2017 we saw the effect of stagflation - while the USD/TL rate increasing (accordingly we know that inflation rate is strongly positively correlatied with the inflation rate (inf rate: 2015-12:5.71%, 2016-12:9.94%, 2017-12:15.47%)) and weak or negative economic growth. 
Besides, there is not a direct correlation between BIST100 index and Trade Deficit Values.

####EMRE

####Analysis of Year by Year Change in Trade Values

```{r}
tableExport <- export_jTableMELTED %>% 
  group_by(Country, year) %>%
  summarise(totalExpV=sum(Values,na.rm=TRUE)) %>%
  arrange(Country, year) %>%
  mutate(ExpValChng = (totalExpV - lag(totalExpV))/lag(totalExpV)*100)

tableImport <- import_jTableMELTED %>% 
  group_by(Country, year) %>%
  summarise(totalImpV=sum(Values,na.rm=TRUE)) %>%
  arrange(Country, year) %>%
  mutate(ImpValChng = (totalImpV - lag(totalImpV))/lag(totalImpV)*100)

table <- full_join(tableExport,tableImport) %>%
  select(Country, year, Export="ExpValChng", Import = "ImpValChng") %>%
  melt(id=c("Country", "year")) %>%
  filter(year!=2013)
```

```{r}
ggplot(table, aes(x=year)) +
  geom_point(aes(y=value, color=Country, shape=variable), size=4) + 
  facet_grid(~variable) +
   labs(title="YoY Total Value Change",
    x="Year",
    y="% Change in  Value",
    shape=c("Trade Type"))
```

```{r}

# Canada and Russia significantly deviates. Investigate these countries in detail with traded products

# First Canada

# Export Analysis

summarytableCanadaEx <- export_jTableMELTED %>%
  filter(Country=="Canada") %>%
  group_by(X__2, year) %>%
  summarise(ExportValue=sum(Values)) %>%
  arrange(X__2, year) %>%
  mutate(ExpPercChngYoY = (ExportValue - lag(ExportValue))/lag(ExportValue)*100) %>%
  mutate(ExpChngYoy=((ExportValue - lag(ExportValue)))) %>%
  filter(year!=2013)

mean(summarytableCanadaEx$ExpChngYoy)
median(summarytableCanadaEx$ExpChngYoy)
sd(summarytableCanadaEx$ExpChngYoy)
quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.25, 0.50, 0.75, 1))
IQR= quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.75)) - quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.25))
altsinir=quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.25)) - IQR
ustsinir=quantile(summarytableCanadaEx$ExpChngYoy, probs = c(0.75)) + IQR

summarytableCanadaEx <- summarytableCanadaEx %>%
  filter(abs(ExpChngYoy)>max(abs(altsinir),ustsinir))

plotCanEx <- ggplot(summarytableCanadaEx, aes(x=substr(X__2, start = 1, stop = 30))) +
  geom_point(aes(y=ExpChngYoy, color=year), size=4) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid (~year) +
  facet_grid(rows = vars(year)) +
  labs(title="YoY Total Value Change with Canada - Export",
       x="Product Group",
       y="Change in Export Value",
       Color=c("Year"))

# Import Analysis

summarytableCanadaIm <- import_jTableMELTED %>%
  filter(Country=="Canada") %>%
  group_by(X__2, year) %>%
  summarise(ImportValue=sum(Values)) %>%
  arrange(X__2, year) %>%
  mutate(ImpPercChngYoY = (ImportValue - lag(ImportValue))/lag(ImportValue)*100) %>%
  mutate(ImpChngYoy=((ImportValue - lag(ImportValue)))) %>%
  filter(year!=2013)

mean(summarytableCanadaIm$ImpChngYoy)
median(summarytableCanadaIm$ImpChngYoy)
sd(summarytableCanadaIm$ImpChngYoy)
quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.25, 0.50, 0.75, 1))
IQR= quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.75)) - quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.25))
altsinir=quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.25)) - IQR
ustsinir=quantile(summarytableCanadaIm$ImpChngYoy, probs = c(0.75)) + IQR

summarytableCanadaIm <- summarytableCanadaIm %>%
  filter(abs(ImpChngYoy)>max(abs(altsinir),ustsinir))

plotCanIm <- ggplot(summarytableCanadaIm, aes(x=substr(X__2, start = 1, stop = 30))) +
  geom_point(aes(y=ImpChngYoy, color=year), size=4) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid (~year) +
  facet_grid(rows = vars(year)) +
  labs(title="YoY Total Value Change Canada - Import",
       x="Product Group",
       y="Change in Import Value",
       Color=c("Year"))
```

```{r}

# Second Russia

# Export Analysis

summarytableRussianEx <- export_jTableMELTED %>%
  filter(Country=="Russian") %>%
  group_by(X__2, year) %>%
  summarise(ExportValue=sum(Values)) %>%
  arrange(X__2, year) %>%
  mutate(ExpPercChngYoY = (ExportValue - lag(ExportValue))/lag(ExportValue)*100) %>%
  mutate(ExpChngYoy=((ExportValue - lag(ExportValue)))) %>%
  filter(year!=2013)

mean(summarytableRussianEx$ExpChngYoy)
median(summarytableRussianEx$ExpChngYoy)
sd(summarytableRussianEx$ExpChngYoy)
quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.25, 0.50, 0.75, 1))
IQR= quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.75)) - quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.25))
altsinir=quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.25)) - IQR
ustsinir=quantile(summarytableRussianEx$ExpChngYoy, probs = c(0.75)) + IQR

summarytableRussianEx <- summarytableRussianEx %>%
  filter(abs(ExpChngYoy)>max(abs(altsinir),ustsinir))

plotRusEx <- ggplot(summarytableRussianEx, aes(x=substr(X__2, start = 1, stop = 30))) +
  geom_point(aes(y=ExpChngYoy, color=year), size=4) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid (~year) +
  facet_grid(rows = vars(year)) +
  labs(title="YoY Total Value Change Russia - Export",
       x="Product Group",
       y="Change in Export Value",
       Color=c("Year"))

# Import Analysis

summarytableRussianIm <- import_jTableMELTED %>%
  filter(Country=="Russian") %>%
  group_by(X__2, year) %>%
  summarise(ImportValue=sum(Values)) %>%
  arrange(X__2, year) %>%
  mutate(ImpPercChngYoY = (ImportValue - lag(ImportValue))/lag(ImportValue)*100) %>%
  mutate(ImpChngYoy=((ImportValue - lag(ImportValue)))) %>%
  filter(year!=2013)

mean(summarytableRussianIm$ImpChngYoy)
median(summarytableRussianIm$ImpChngYoy)
sd(summarytableRussianIm$ImpChngYoy)
quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.25, 0.50, 0.75, 1))
IQR= quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.75)) - quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.25))
altsinir=quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.25)) - IQR
ustsinir=quantile(summarytableRussianIm$ImpChngYoy, probs = c(0.75)) + IQR

summarytableRussianIm <- summarytableRussianIm %>%
  filter(abs(ImpChngYoy)>max(abs(altsinir),ustsinir))

plotRusIm <- ggplot(summarytableRussianIm, aes(x=substr(X__2, start = 1, stop = 30))) +
  geom_point(aes(y=ImpChngYoy, color=year), size=4) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_grid (~year) +
  facet_grid(rows = vars(year)) +
  labs(title="YoY Total Value Change Russia - Import",
       x="Product Group",
       y="Change in Import Value",
       Color=c("Year"))
```

```{r}

#install.packages("gridExtra")

library(gridExtra)

grid.arrange(plotCanEx, plotCanIm,  nrow = 1)
grid.arrange(plotRusEx, plotRusIm,  nrow = 2)
